---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: yes
---


## Loading and preprocessing the data

The activity monitoring data is unzipped and loaded. The summary is shown below.

```{r loading, cache=TRUE}
unzip("act.zip")
activity <- read.csv("activity.csv")
str(activity)
```

## What is mean total number of steps taken per day?

```{r dailay steps}
d_steps <- with(activity, tapply(steps, date, sum, na.rm = TRUE))
d_mean <- as.integer(mean(d_steps, na.rm = TRUE))
d_median <- as.integer(median(d_steps, na.rm = TRUE))
hist(d_steps, breaks = 10, xlab = "Steps", ylab = "days",
     main = "Histogram of the Total Steps Each Day")
```

The mean number of steps taken per day is `r d_mean` and the median number is `r d_median`.

## What is the average daily activity pattern?

The average daily activity pattern is shown as the number of steps taken over time series
```{r daily pattern}
t_steps <- with(activity, tapply(steps, interval, mean, na.rm = TRUE))
barplot(t_steps, main = "Average Daily Activity Pattern")
mt <- activity[which.max(t_steps), 3]
```

The interval `r mt` contains the maximum number of steps.

## Imputing missing values

The missing values are found by
```{r missing values}
row_na <- is.na(activity$steps)
mv <- sum(row_na)
```

There are `r mv` missing values in the original data set.
The missing values can be filled with average steps taken on each interval over the testing dates.

```{r fill in missing values}
activity$steps2 <- replace(activity[ , 1], which(row_na), t_steps)
d_steps2 <- with(activity, tapply(steps2, date, sum, na.rm = TRUE))
hist(d_steps2, breaks = 10, xlab = "Steps", ylab = "days",
     main = "Histogram of the Total Steps Each Day after NA filled")
mean_steps2 <- as.integer(mean(d_steps2, na.rm = TRUE))
median_steps2 <- as.integer(median(d_steps2, na.rm = TRUE))
```

So after the missing values filled, the mean number of steps taken each day is `r mean_steps2` and the median number is `r median_steps2`. Both numbers are higher than the original ones because the missing values were counted as 0 when calculating the total number of daily steps.

## Are there differences in activity patterns between weekdays and weekends?

```{r compare weekdays}
activity$day <- weekdays(as.Date(activity$date))
wdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
ends <- c("Saturday", "Sunday")
activity$wk <- activity$day
for(i in 1:nrow(activity)) {
if(activity$day[i] %in% ends) activity$wk[i] <- "weekends"
else activity$wk[i] <- "weekdays"
}
t_steps2 <- with(activity, tapply(steps2, list(interval, wk), mean, na.rm = TRUE))
library(reshape2)
library(ggplot2)
t_steps3 <- melt(t_steps2)
names(t_steps3) <- c("interval", "weekday", "steps")
qplot(interval, steps, data = t_steps3, facets = weekday~., 
      geom = "line", color = weekday)
```

In average, the activity patterns of weekdays and weekends are different in 

1. The active time starts earlier in weekdays than in weekends. In weekdays, the activity starts at ~5:00 while it starts after ~7:00 in weekends.

2. The rest time also starts earlier in weekdays than in weekends. In weekdays, the activity stops after ~20:00 while it stops after ~21:00 in weekends.  

3. Both patterns in weekdays and weekends contain multiple peaks while the pattern of weekdays contains an obvious high peak above 200 steps at ~8:30.
